draft 1

### 2015-10-20 boost_tree.rmd
more boosting

```{r, load library and packages, include= FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')

@@

all8a$b.rpdol<- all8a$return.pdol 
all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 0, 1)

@@

all9a<- all8a
```

@@

```{r, echo=FALSE}
formula = "b.rpdol~.- mlsto - return.pdol-Year-code.client-code.contact"

errors<- double(4)
shrink.seq<-c(.001,.005,.01,.05)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees=5000,
                 shrinkage=shrink.seq[i],interaction.depth=3)
        pred<- predict(fit, test, n.tree=5000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=shrink.seq, "error" = errors[1:4])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

@@

Try caret method of tuning

```{r, echo=FALSE, include=FALSE}
# must turn outcome into two level factor
# all9a$f.rpdol<- as.factor(all9a$b.rpdol)
# set.seed(100)
# sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
# train<- all9a[sample,]
# test<- all9a[-sample,]
# 
# # set seeds
# set.seed(123)
# seeds <- vector(mode = "list", length = 51)
# for(i in 1:50) seeds[[i]] <- sample.int(1000, 22)
# ## For the last model:
# seeds[[51]] <- sample.int(1000, 1)
# 
# # first tune shrinkage and number of trees
# cv.Control <- trainControl(method = "cv",
#                            number = 10,
#                            seeds= seeds)
# gbmGrid <- expand.grid(shrinkage = c(.001,.005),
#                        n.trees = c(2000), interaction.depth = 2,
#                        n.minobsinnode = 10)
# set.seed(2)
# gbmFit <- train(f.rpdol~.- mlsto - return.pdol-Year, data= train,
#                  method = "gbm", trControl = cv.Control, verbose = FALSE,
#                  bag.fraction = 0.5, tuneGrid = gbmGrid,
#                 metric = 'kappa')
# 
# plot(gbmFit)
```

Caret package is not working, tune manually:

@@

#shrinkage should be 0.001, now find number of trees
errors<- double(5)
ntree.seq<-c(4000,8000,10000,12000,14000)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= ntree.seq[i],
                 shrinkage=0.001,interaction.depth=3)
        pred<- predict(fit, test, n.tree=ntree.seq[i], type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=ntree.seq, "error" = errors[1:5])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

#10000 trees seems best?

@@

#shrinkage should be 0.001, 10000 trees, now find interaction.depth
errors<- double(5)
int.seq<-c(1,2,4,6,8)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=int.seq[i])
        pred<- predict(fit, test, n.tree=10000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=int.seq, "error" = errors[1:5])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

## use interaction.depth = 4?

#------------------------------------------------
# now tune min obs in node
errors<- double(4)
obs.seq<-c(10,20,40,50)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 10000,
                 shrinkage=0.001,interaction.depth=4,
                 n.minobsinnode = obs.seq[i])
        pred<- predict(fit, test, n.tree=10000, type="response")
        pred<- ifelse(pred>=0.5,1,0)
        temp<- ifelse(pred == test$b.rpdol, 0,1)
        errors[i]<- sum(temp)/nrow(test)

        #confusion matrix
        temp2<- data.frame('pred' = pred, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
errors

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=obs.seq, "error" = errors[1:4])
ggplot(shrink, aes(x=shrinkage, y=error)) + geom_line() + geom_point()+ geom_line()

>>>>

```{r, load library and packages, include= FALSE}
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('ggplot2',lib='C:/Progra~1/R/R-3.2.1/library')
library('gbm', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('caret', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ROCR',lib='C:/Progra~1/R/R-3.2.2/library')

@@

all8a$b.rpdol<- all8a$return.pdol 
all8a$b.rpdol<- ifelse(all8a$return.pdol<=0, 1, 0)

@@

all9a<- all8a
write.csv(all9a, 'C:/Users/n9232371/Documents/Consultbusiness/data/all9a.csv')

@@

Try deleting code.client variable

#tune which variables should be included!!
vars.excl = c("JD.Second", "Business", "JD.Second-Business","")

errors<- double(4)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
auc<- double(4)
colours <- c("red","green","blue", "pink")

for(i in 1:4){
        if(vars.excl[i] %in% ""){
              formula<- "b.rpdol~.- mlsto - return.pdol-Year - code.contact - code.client"  
        } else{
        formula<- paste("b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact",
                        "-", vars.excl[i], sep=" ")
        }

        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees=5000,
                 shrinkage=0.001,
                 interaction.depth=3)
        pred<- predict(fit, test, n.tree=5000, type="response")
        errors[i]<- sum(temp)/nrow(test)

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres=T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred1' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred1, temp2$ans))
        cat(i," ")
}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=shrink.seq, "auc" = auc[1:4])
ggplot(shrink, aes(x=shrinkage, y= auc)) + geom_line() + geom_point()+ geom_line()

#best to not have JD.Second OR Business!!!  :(
```

Tune shrinkage value:
```{r, echo=FALSE}

formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business"

errors<- double(4)
shrink.seq<-c(0.0005, .001,.005, 0.01)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
auc<- double(4)
colours <- c("red", "yellow", "green", "blue")

for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees=5000,
                 shrinkage=shrink.seq[i],interaction.depth=3)
        pred<- predict(fit, test, n.tree=5000, type="response")
        errors[i]<- sum(temp)/nrow(test)

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres=T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred1' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred1, temp2$ans))
        cat(i," ")
}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=shrink.seq, "auc" = auc[1:4])
ggplot(shrink, aes(x=shrinkage, y= auc)) + geom_line() + geom_point()+ geom_line()

#best case auc = 0.756 with shrinkage = 0.001

@@

@@

#shrinkage should be 0.001, now find number of trees
errors<- double(5)
ntree.seq<-c(2000,3000,4000,6000,8000)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
auc<- double(5)
colours <- c("red", "yellow", "green", "blue", "purple")

for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= ntree.seq[i],
                 shrinkage=0.001,interaction.depth=3)
        pred<- predict(fit, test, n.tree=ntree.seq[i], type="response")
        errors[i]<- sum(temp)/nrow(test)

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres=T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=ntree.seq, "auc" = auc[1:5])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

#4000 trees seems best?

@@

#shrinkage should be 0.001, 4000 trees, now find interaction.depth
errors<- double(5)
int.seq<-c(1,2,3,4,6)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
auc<- double(5)
colours <- c("red", "yellow", "green", "blue", "purple")

for(i in 1:5){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 4000,
                 shrinkage=0.001,interaction.depth=int.seq[i])
        pred<- predict(fit, test, n.tree=4000, type="response")
        errors[i]<- sum(temp)/nrow(test)

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres=T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE, print.thres=T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=int.seq, "auc" = auc[1:5])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

## use interaction.depth = 3

#------------------------------------------------
# now tune min obs in node

errors<- double(4)
obs.seq<-c(10,20,40,50)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)))
auc<- double(4)
colours <- c("red", "yellow", "green", "blue")

for(i in 1:4){
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 4000,
                 shrinkage=0.001,interaction.depth=3,
                 n.minobsinnode = obs.seq[i])
        pred<- predict(fit, test, n.tree=4000, type="response")

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres = T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")
}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=obs.seq, "auc" = auc[1:4])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

#go with 10

So paramters are ....

    * shrinkage = 0.001
    * exclude vars = JD.Second, Business, code.client, Year, code.contact
    * n.trees = 4000
    * interaction.depth = 3
    * minobsinnode = 10

to obtain AUC = 0.785

But can we improve the AUC by bringing back some of the missing variables in a smart way??

For example, bring back code.client, but only if clients have more than say 10 jobs?

#new variable for clients with more than 10 jobs
#list of clients with more than 10 jobs
all9b<- all9a
client.count<- ddply(all9b, .(code.client), nrow)
client.count<- arrange(client.count, -V1)

client.count2<- client.count[client.count$V1>20 & !is.na(client.count$code.client),]

all9b$code.client<- as.character(all9b$code.client)
all9b$reduc.client<- ifelse(all9b$code.client %in% client.count2$code.client,
                            all9b$code.client,
                            NA)
all9b$code.client<- as.factor(all9b$code.client)
all9b$reduc.client<- as.factor(all9b$reduc.client)

#see if ROC improves

formula<- "b.rpdol~.- mlsto - return.pdol-Year-code.client - code.contact-JD.Second - Business"

client.lim <- c(20,30,40,60,70)
conf.matrix<- list('1' = data.frame(c(0,0), c(0,0)), '2' = data.frame(c(0,0), c(0,0)),
                   '3' = data.frame(c(0,0), c(0,0)), '4' = data.frame(c(0,0), c(0,0)),
                   '5' = data.frame(c(0,0), c(0,0)))
auc<- double(5)
colours <- c("red", "yellow", "green", "blue", "purple")

for(i in 1:5){

        client.count2<- client.count[client.count$V1> client.lim[i] &
                                             !is.na(client.count$code.client),]
        all9b$code.client<- as.character(all9b$code.client)
        all9b$reduc.client<- ifelse(all9b$code.client %in% client.count2$code.client,
                            all9b$code.client,
                            NA)
        all9b$code.client<- as.factor(all9b$code.client)
        all9b$reduc.client<- as.factor(all9b$reduc.client)

        #see if ROC improves

        #redo train and test
        set.seed = 300
        sample<- sample(1:nrow(all9a), 2/3*nrow(all9a), replace=F)
        train<- all9a[sample,]
        test<- all9a[-sample,]

        #make boosted tree
        fit<-gbm(as.formula(formula), 
                 data=train, distribution="bernoulli", n.trees= 4000,
                 shrinkage=0.001,interaction.depth=3,
                 n.minobsinnode = 10)
        pred<- predict(fit, test, n.tree=4000, type="response")

        #pROC
        pred.p<- roc(test$b.rpdol, pred)

        if(i==1){
                plot.roc(pred.p, col= colours[i], print.thres = T)

        } else{
                plot.roc(pred.p, col= colours[i], add=TRUE , print.thres = T)
        }

        auc[i]<- auc(pred.p)

        #confusion matrix
        conf.mat<- coords(pred.p, "best")
        pred1<- ifelse(pred>= conf.mat[1], 1, 0)
        temp2<- data.frame('pred' = pred1, 'ans' = test$b.rpdol)
        conf.matrix[[i]] = as.data.frame.matrix(table(temp2$pred, temp2$ans))
        cat(i," ")

}

conf.matrix
auc

#create new data frame 'shrink' to place trees and errors vectors together, then plot:
shrink<- data.frame("shrinkage"=client.lim, "auc" = auc[1:5])
ggplot(shrink, aes(x=shrinkage, y=auc)) + geom_line() + geom_point()+ geom_line()

#best I can do is 0.783
```

Make a function to do the same with code.contact, JD.Second, and Business????

### 2015-10-20 finalwriteup/draft.rmd
more boosting

>>>>

# Chapter ? Method

## Regression
    * tried predicting return per dollar as a numeric outcome.
        * proved very difficult. Find error rate compared to just guessing the mean. Not far off.

##Bayesian Network

* could not handle _any_ numeric variables, had to discretise all numeric vars
    * used normalised numeric variables
    * I did this by making a hierarchical dendrogram to visualise the clusters
    * then I chose the number of clusters, drew the red rectangle
    * summarised the clusters to find max and min values for the variable within each cluster. Rounded this value up
    * manually discretised variable using rounded values

* first trialled complete cases of number of users, Discipline, percent pro, Business, client code, job detail, majority position, timespan, invoiced amount, percent hours by majority position, and return per dollar. This gave 907 complete cases to work with.

*using Genie Smile.
   * turned return per dollar into profit/loss. This gave 182 loss cases and 725 profit cases. A good distribution for working with bayesian networks. Ie not too biased
   * a good, simple problem to start with

* establish a baseline. Using ROC, area under the curve method
    * to test our classifiers we need to beat random chance. Ie AUC = 0.5
    * will also establish the AUC for using one variable only! See table below:
       *Note all using 10 fold cross validation

Single Variable  | AUC from ROC chart | Num losses correct
---------------- | ------------------ | ------------------
Random chance    | 0.5                | ?                 
percent pro      | 0.618              | 0
percent maj pos  | 0.616              | 0
number of users  | 0.661              | 0
timespan         | 0.600              | 0
amount invoiced  | 0.535              | 0
Discipline       | 0.514              | 0
Majority pos     | 0.605              | 3
Job detail       | 0.605              | 1
client code      | 0.567              | 25
client Business  | 0.524              | 5

* try using the following variables:

```{r, echo=FALSE, include=FALSE}

read.csv(all8c,'C:/Users/n9232371/Documents/Consultbusiness/data/all8c.csv' )

vars<- names(all8c)
vars<- vars[-c("Business", "JD_Second", "code_client")]

```

* number of users
* Discipline
* percent hours by a professional
* position that completed the majority of hours
* timespan
* amount invoiced
* percent of hours completed by majority position

The results are as follows:

AUC = 0.705
Number of losses predicted correctly: 124

This is our 'base' model. Now try adding one complex variable at a time:

Added Variable   | AUC from ROC chart | Num losses correct (tp) | Num losses wrong (fp)
---------------- | ------------------ | ----------------------- | --------------------
none/base        | 0.710              | 122                     | 272
client code      | 0.624              | 171                     | 528
Busines          | 0.667              | 144                     | 381
job detail       | 0.653              | 150                     | 438

## Boosted Trees

### Advantages

    * Wonderful in that NA's don't interfere with prediciton process
        * explain..
        * can comfortably include the entire dataset without having to worry about imputing

tuned which variables to include:
    * exclude JD.Second, Business, code.client, code.contact
    * tuned parameters:
        * shrinkage = 0.001
        * n.trees = 4000
        * interaction.depth = 3
        * min.nobs = 10
    * manage to get AUC up to 0.785!!

can we improve this by bringing huge variables back in a smarter way?
